\large \name: Combining spectra or images by forward modeling

### 2024-03-18 paper/main.tex
new figure locations
changes for AAS submisssion

\documentclass[modern]{aastex631}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}

@@

According to these rules we made two input data sets.
One is poorly sampled, with $M=171$ spectral pixels in the input data separated by $2/R$ in $x$ (natural logarithm of wavelength), and signal-to-noise per pixel of 18 in the continuum.
The other is well sampled, with $M=340$ spectral pixels in the input data separated by $1/R$ in $x$, and signal-to-noise per pixel of 12 in the continuum.
In each case there are $N=8$ epochs.
The two complete data sets are shown in \figurename~\ref{fig:data1} and \figurename~\ref{fig:data2}.
\begin{figure}[t!]
    \begin{mdframed}\begin{center}
    \includegraphics[width=1.3\figurewidth]{notebooks/data1.pdf}\\
    \includegraphics[width=\figurewidth]{notebooks/datazoom.pdf}
    \end{center}
    \caption{The top 8 panels show the $N=8$ epochs of the poorly sampled input data, with $N=8$, $M=171$, and signal-to-noise per pixel of 18 in the continuum. The bad pixels are indicated with grey bars. The bottom 2 panels show a zoom in on one of the epochs, along with the true spectrum used to generate the data. The true spectrum is shown at zero Doppler shift, while each epoch spectrum is at a finite Doppler shift $\Delta x_i$ relative to the spectrograph pixel grid.\label{fig:data1}}
    \end{mdframed}
\end{figure}
\begin{figure}[t!]
    \begin{mdframed}\begin{center}
    \includegraphics[width=1.3\figurewidth]{notebooks/data2.pdf}
    \end{center}
    \caption{Similar to the top panels of \figurename~\ref{fig:data1} except showing the well sampled input data with $N=8$, $M=340$, and signal-to-noise per pixel of 12 in the continuum.\label{fig:data2}}
    \end{mdframed}
\end{figure}

@@

\begin{figure}[t!]
    \begin{mdframed}\begin{center}
    \includegraphics[width=\figurewidth]{notebooks/forward1.pdf}\\
    \includegraphics[width=\figurewidth]{notebooks/forward2.pdf}
    \end{center}
    \caption{The top two panels show the result of running \name{} on the poorly sampled input data, plus the residual away from the truth. The bottom two panels show the same but for the well sampled input data. In each panel, the red line shows the true model used to generate the data.\label{fig:forward}}
    \end{mdframed}
\end{figure}
The combined spectrum resulting from the forward model described in \secref{sec:method} is shown in \figurename~\ref{fig:forward}.
It is compared to the true spectral expectation model that was used to generate the data.
Also shown in \figurename~\ref{fig:forward} is the residuals, \name{} minus truth.
The residuals look stationary, approximately Gaussian, and uncorrelated (more on this below).

@@

\begin{figure}[t!]
    \begin{mdframed}\begin{center}
    \includegraphics[width=\figurewidth]{notebooks/datazoom_interpolated.pdf}
    \end{center}
    \caption{As part of Standard Practice(tm), it is necessary to interpolate the individual epoch spectra before averaging them. This shows one epoch of the poorly sampled data, interpolated to the output $x_\star$ pixel grid in the rest frame. The interpolated spectrum is compared to the true spectral expectation employed to generate it. Note the ringing induced by the interpolation. Also shown is the censored data, near an interpolated bad pixel and at the edge of the domain. Compare to the bottom panels of \figurename~\ref{fig:data1}.\label{fig:interpolated}}
    \end{mdframed}
\end{figure}
An example of a poorly-sampled spectrum interpolated to the rest frame is shown in \figurename~\ref{fig:interpolated}.
The interpolation introduces substantial ringing artifacts in the spectrum (compare to \figurename~\ref{fig:data1}).
It also grows the bad-pixel mask around the bad pixel.

\begin{figure}[t!]
    \begin{mdframed}\begin{center}
    \includegraphics[width=\figurewidth]{notebooks/standard1.pdf}\\
    \includegraphics[width=\figurewidth]{notebooks/standard2.pdf}
    \end{center}
    \caption{Same as \figurename~\ref{fig:forward} but for the output of Standard Practice(tm). Note that in the poorly sampled case, the residuals appear to be spatially correlated, and the residuals appear to be larger near strong spectral features; more about these issues in the text and in \figurename~\ref{fig:noise}.\label{fig:standard}}
    \end{mdframed}
\end{figure}
We show the final result of Standard Practice(tm)---interpolating the input data and then averaging the interpolations---for the two data sets, in \figurename~\ref{fig:standard}.
We also show the comparison with the true spectral expectation model that was employed to make the fake data.
The residuals look correlated (there is ringing which mimics weak spectral features), and the residuals are larger near strong spectral features, especially in the case of the poorly sampled input data.

\paragraph{Noise and noise covariances}
We repeated the above experiments to empirically measure the noise covariances, and to estimate how correlated neighboring pixels are in the output spectrum. For the poorly sampled case and the well sampled case, we generated 64 multi-epoch data sets and computed the mean spectrum using \name{} and Standard Practice(tm). We then computed the empirical covariance in the output spectrum minus the true spectrum, which are shown in \figurename~\ref{fig:noise}.
In each case, Standard Practice(tm) has a lower noise variance at zero lag, but shows spatially correlated noise over many pixels.
\name{}, in contrast, shows no pixel-to-pixel noise covariances, even when the input data are ill sampled.
\begin{figure}[t!]
    \begin{mdframed}\begin{center}
    \includegraphics[width=\figurewidth]{notebooks/noise.pdf}
    \end{center}
    \caption{The empirical noise variance in the output (combined) spectra, and pixel-to-pixel covariances, as a function of pixel lag (pixel offset). The variances were estimated by performing the experiments repeated 64 times, with the same spectral expectations and sampling but unique noise draws. \name{} has slightly larger variance per pixel at zero lag, but vanishing pixel-to-pixel correlations. In contrast, the Standard Practice(tm) leads to correlated pixel values, with correlations extending out to many pixels.\label{fig:noise}}
    \end{mdframed}
\end{figure}

@@

\paragraph{Acknowledgements}
It is a pleasure to thank
Mike Blanton (NYU),
Matt Daunt (NYU),
Dylan Green (UC Irvine),
Jon Holtzman (NMSU), 
Dustin Lang (Perimeter),
Adrian Price-Whelan (Flatiron),
and the Astronomical Data Group at the Flatiron Institute
for valuable discussions and input.
The Flatiron Institute is a division of the Simons Foundation.
This research was supported in part by the Australian Research Council Centre of Excellence for All Sky Astrophysics in 3 Dimensions (ASTRO 3D), through project number CE170100013.

>>>>

\documentclass[modern, linenumbers]{aastex631}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}

@@

According to these rules we made two input data sets.
One is poorly sampled, with $M=171$ spectral pixels in the input data separated by $2/R$ in $x$ (natural logarithm of wavelength), and signal-to-noise per pixel of 18 in the continuum.
The other is well sampled, with $M=340$ spectral pixels in the input data separated by $1/R$ in $x$, and signal-to-noise per pixel of 12 in the continuum.
In each case there are $N=8$ epochs.
The two complete data sets are shown in \figurename~\ref{fig:data1} and \figurename~\ref{fig:data2}.
\begin{figure}[t!]
    \begin{mdframed}\begin{center}
    \includegraphics[width=1.3\figurewidth]{data1.pdf}\\
    \includegraphics[width=\figurewidth]{datazoom.pdf}
    \end{center}
    \caption{The top 8 panels show the $N=8$ epochs of the poorly sampled input data, with $N=8$, $M=171$, and signal-to-noise per pixel of 18 in the continuum. The bad pixels are indicated with grey bars. The bottom 2 panels show a zoom in on one of the epochs, along with the true spectrum used to generate the data. The true spectrum is shown at zero Doppler shift, while each epoch spectrum is at a finite Doppler shift $\Delta x_i$ relative to the spectrograph pixel grid.\label{fig:data1}}
    \end{mdframed}
\end{figure}
\begin{figure}[t!]
    \begin{mdframed}\begin{center}
    \includegraphics[width=1.3\figurewidth]{data2.pdf}
    \end{center}
    \caption{Similar to the top panels of \figurename~\ref{fig:data1} except showing the well sampled input data with $N=8$, $M=340$, and signal-to-noise per pixel of 12 in the continuum.\label{fig:data2}}
    \end{mdframed}
\end{figure}

@@

\begin{figure}[t!]
    \begin{mdframed}\begin{center}
    \includegraphics[width=\figurewidth]{forward1.pdf}\\
    \includegraphics[width=\figurewidth]{forward2.pdf}
    \end{center}
    \caption{The top two panels show the result of running \name{} on the poorly sampled input data, plus the residual away from the truth. The bottom two panels show the same but for the well sampled input data. In each panel, the red line shows the true model used to generate the data.\label{fig:forward}}
    \end{mdframed}
\end{figure}
The combined spectrum resulting from the forward model described in \secref{sec:method} is shown in \figurename~\ref{fig:forward}.
It is compared to the true spectral expectation model that was used to generate the data.
Also shown in \figurename~\ref{fig:forward} is the residuals, \name{} minus truth.
The residuals look stationary, approximately Gaussian, and uncorrelated (more on this below).

@@

\begin{figure}[t!]
    \begin{mdframed}\begin{center}
    \includegraphics[width=\figurewidth]{datazoom_interpolated.pdf}
    \end{center}
    \caption{As part of Standard Practice(tm), it is necessary to interpolate the individual epoch spectra before averaging them. This shows one epoch of the poorly sampled data, interpolated to the output $x_\star$ pixel grid in the rest frame. The interpolated spectrum is compared to the true spectral expectation employed to generate it. Note the ringing induced by the interpolation. Also shown is the censored data, near an interpolated bad pixel and at the edge of the domain. Compare to the bottom panels of \figurename~\ref{fig:data1}.\label{fig:interpolated}}
    \end{mdframed}
\end{figure}
An example of a poorly-sampled spectrum interpolated to the rest frame is shown in \figurename~\ref{fig:interpolated}.
The interpolation introduces substantial ringing artifacts in the spectrum (compare to \figurename~\ref{fig:data1}).
It also grows the bad-pixel mask around the bad pixel.

\begin{figure}[t!]
    \begin{mdframed}\begin{center}
    \includegraphics[width=\figurewidth]{standard1.pdf}\\
    \includegraphics[width=\figurewidth]{standard2.pdf}
    \end{center}
    \caption{Same as \figurename~\ref{fig:forward} but for the output of Standard Practice(tm). Note that in the poorly sampled case, the residuals appear to be spatially correlated, and the residuals appear to be larger near strong spectral features; more about these issues in the text and in \figurename~\ref{fig:noise}.\label{fig:standard}}
    \end{mdframed}
\end{figure}
We show the final result of Standard Practice(tm)---interpolating the input data and then averaging the interpolations---for the two data sets, in \figurename~\ref{fig:standard}.
We also show the comparison with the true spectral expectation model that was employed to make the fake data.
The residuals look correlated (there is ringing which mimics weak spectral features), and the residuals are larger near strong spectral features, especially in the case of the poorly sampled input data.

\paragraph{Noise and noise covariances}
We repeated the above experiments to empirically measure the noise covariances, and to estimate how correlated neighboring pixels are in the output spectrum. For the poorly sampled case and the well sampled case, we generated 64 multi-epoch data sets and computed the mean spectrum using \name{} and Standard Practice(tm). We then computed the empirical covariance in the output spectrum minus the true spectrum, which are shown in \figurename~\ref{fig:noise}.
In each case, Standard Practice(tm) has a lower noise variance at zero lag, but shows spatially correlated noise over many pixels.
\name{}, in contrast, shows no pixel-to-pixel noise covariances, even when the input data are ill sampled.
\begin{figure}[t!]
    \begin{mdframed}\begin{center}
    \includegraphics[width=\figurewidth]{noise.pdf}
    \end{center}
    \caption{The empirical noise variance in the output (combined) spectra, and pixel-to-pixel covariances, as a function of pixel lag (pixel offset). The variances were estimated by performing the experiments repeated 64 times, with the same spectral expectations and sampling but unique noise draws. \name{} has slightly larger variance per pixel at zero lag, but vanishing pixel-to-pixel correlations. In contrast, the Standard Practice(tm) leads to correlated pixel values, with correlations extending out to many pixels.\label{fig:noise}}
    \end{mdframed}
\end{figure}

@@

\begin{acknowledgments}
It is a pleasure to thank
Mike Blanton (NYU),
Matt Daunt (NYU),
Dylan Green (UC Irvine),
Jon Holtzman (NMSU), 
Dustin Lang (Perimeter),
Adrian Price-Whelan (Flatiron),
and the Astronomical Data Group at the Flatiron Institute
for valuable discussions and input.
The Flatiron Institute is a division of the Simons Foundation.
This research was supported in part by the Australian Research Council Centre of Excellence for All Sky Astrophysics in 3 Dimensions (ASTRO 3D), through project number CE170100013.
\end{acknowledgments}
